Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-19 doc/manuscript/translocation_pnas_SI.tex
Adjust manuscript formatting

In the comparison of individual outlier variants between naive and
recovering populations, we detected consistent allele frequency changes
across all genetic clusters for 1 of 7 of the associated genes (RIN3;
Fig. 1). Nonetheless, for the other 6 genes
we still found a statistically significant signal of selection in 2 of
the 3 genetic clusters (containing populations 6, 7, 8, and 9;
Fig. 1, Figure~\ref{fig-selectionresults} A
SI). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe parallel
evolutionary changes following Bd exposure and are therefore interesting
candidates for future studies.

@@

The baseline parameter values for the model and how they were estimated
are given in Table~\ref{tbl-param_values} : SI. Parameters
\(\sigma_{A_T}\) and \(\sigma_{A_R}\) were extracted directly from our
CMR model described previously in this study. For populations where we
had a sufficient number of PIT-tagged, naturally-recruited adults, we
observed that \(\sigma_{A_T}\) and \(\sigma_{A_R}\) could be notably
different, with \(\sigma_{A_R} > \sigma_{A_T}\)
(Figure~\ref{fig-compare_surv_probs} : SI). For populations lacking
sufficient numbers of naturally-recruited adults, we were unable to
directly estimate \(\sigma_{A_R}\), and instead set
\(\sigma_{A_R} = \sigma_{A_T}\).

@@

\caption{\label{fig-selectionresults}Evidence for selection of
individual variants in recovering MYL frog populations. (A) PCA
calculated from binary SNPs showing the genomic relationship of samples.
Numeric labels and colors match those in
Fig. 1 A. (B) qqplot showing observed and
expected p-values for 148,307 SNPs and INDELS as calculated in GEMMA.
Dashed line shows p-value that identifies outliers. (C) Manhattan plot
showing p-values for each SNP as calculated by GEMMA. SNPs are sorted by
genomic position and contigs are sorted by size. Red line shows p-value
that identifies outliers. Outlier SNPs above this threshold are
highlighted and labeled. Bold labels indicate at least one
non-synonymous SNP is present in that gene.}

@@

\caption{\label{fig-synteny-plot}Synteny plot showing conserved gene
order in \emph{Xenopus tropicalis} and \emph{Rana mucosa} for the
outlier region containing MHC Class I Classical and MHC Class III gene
regions. The plot was created with SimpleSynteny \citep{veltri2016}
using \emph{Xenopus tropicalis} Chromosome 8 (NC\_030684.2, genbank
accession GCA\_000004195.4) and \emph{Rana mucosa} Contig19. Asterices
indicate the location of SNP outliers in TCF19 and VARS1 genes. Gap
sizes for each contig representation are labeled.}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/map_translocation_points.png}

@@

\caption{\label{fig-yosemap}Map showing the locations of translocated
and donor MYL frog populations in Yosemite National Park (park boundary
indicated by black polygon). Symbol shapes indicate the donor population
used for each translocation site. To obscure the exact locations of
populations, random noise was added to all point coordinates. Inset map
shows the location of Yosemite within California. In both maps,
elevation is indicated by the colored hillshade layer (dark green =
lowest elevation, white = highest elevation). }

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/mcmc_areas_m2b.png}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/bdload_beforeafter.png}

@@

\caption{\label{fig-bdload-beforeafter}Evidence that frogs translocated
from recovering populations can maintain the benefits of
resistance/tolerance in non-natal habitats. For frogs translocated to
each of the 12 recipient sites, Bd loads for the period immediately
prior to translocation versus during the 1-year period after
translocation. Bd loads are expressed as the number of ITS1 copies per
skin swab, as estimated by qPCR of the Bd ITS1 region. Loads indicative
of severe disease are \textgreater{} 5.8 ITS copies (on a
log\textsubscript{10} scale; see SI text for details). Box plot details
are as in Fig. 4.}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/mcmc_areas_m1d.png}

@@

\caption{\label{fig-survival-postdens}Evidence that Bd load in not an
important predictor of post-translocation frog survival. For the best
model identifying predictors of 1-year frog survival following
translocation, estimated posterior density curves and shaded 95\%
uncertainty intervals for the intercept and all predictor variables. In
the Bayesian framework in which the model was developed, variables are
considered important predictors if the associated uncertainty interval
does not overlap zero (indicated by the dashed red line).}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/compare_surv_probs.jpg}

@@

\caption{\label{fig-compare_surv_probs}Evidence that frog
resistance/tolerance in non-natal habitats is maintained across
generations. Comparison of the average yearly adult survival
probabilities for adults translocated to each of three sites versus
adults that were naturally recruited at each site (as a result of
reproduction by translocated frogs). In contrast to
Fig. 3, these are not survival
probabilities from the first year following translocation, but instead
represent averaged survival probabilities across multiple years and
cohort. Points are median estimates and error bars give the 95\%
uncertainty intervals around the estimates, accounting for yearly
variation in survival probabilities. The large error bars for 70550
reflect the large within and among individual variability in yearly
survival probability at this site (as also reflected in one-year
survival probabilities in Fig. 3). All
estimates are derived from the mrmr package.}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/pop_viability_figures_for_supp.jpg}

@@

\hypertarget{tbl-survival-earlylate}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0654}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2056}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2430}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2056}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2804}}@{}}
\caption{\label{tbl-survival-earlylate}Association between the
proportion of populations translocated early versus late in the study
period (\textless{} 2013 or \textgreater= 2013, respectively) and
probability of survival (\textless{} 0.5 or \textgreater= 0.5). Based on
the viability analysis, survival probabilities \textless{} 0.5 and
\textgreater= 0.5 produced 50-year extinction probabilities of 1 and
\textless{} 0.5, respectively.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Period
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Survival probability
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Number of translocations
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Total translocations
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Proportion of translocations
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Period
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Survival probability
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Number of translocations
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Total translocations
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Proportion of translocations
\end{minipage} \\
\midrule()
\endhead
early & \textless{} 0.5 & 4 & 5 & 0.800 \\
early & \textgreater= 0.5 & 1 & 5 & 0.200 \\
late & \textless{} 0.5 & 2 & 7 & 0.286 \\
late & \textgreater= 0.5 & 5 & 7 & 0.714 \\
\bottomrule()
\end{longtable}

>>>>

In the comparison of individual outlier variants between naive and
recovering populations, we detected consistent allele frequency changes
across all genetic clusters for 1 of 7 of the associated genes (RIN3;
Fig. 1). However, for the other 6 genes we
still found a statistically significant signal of selection in 2 of the
3 genetic clusters (containing populations 6--9;
Fig. 1, Figure~\ref{fig-selectionresults}
A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe parallel
evolutionary changes following Bd exposure and are therefore interesting
candidates for future studies.

@@

The baseline parameter values for the model and how they were estimated
are given in Table~\ref{tbl-param_values}. Parameters \(\sigma_{A_T}\)
and \(\sigma_{A_R}\) were extracted directly from our CMR model
described previously in this study. For populations where we had a
sufficient number of PIT-tagged, naturally-recruited adults, we observed
that \(\sigma_{A_T}\) and \(\sigma_{A_R}\) could be notably different,
with \(\sigma_{A_R} > \sigma_{A_T}\)
(Figure~\ref{fig-compare_surv_probs}). For populations lacking
sufficient numbers of naturally-recruited adults, we were unable to
directly estimate \(\sigma_{A_R}\), and instead set
\(\sigma_{A_R} = \sigma_{A_T}\).

@@

\caption{\label{fig-selectionresults}Evidence for selection on
individual variants in recovering MYL frog populations. (A) PCA
calculated from binary SNPs showing the genomic relationship of samples.
Numeric labels and colors match those in
Fig. 1 A. Populations 1-7 are \emph{R.
sierrae} and populations 8 and 9 are \emph{R. muscosa}. (B) qqplot
showing observed and expected p-values for 148,307 SNPs and INDELS.
Dashed line shows p-value that identifies outliers. (C) Manhattan plot
showing p-value for each SNP. SNPs are sorted by genomic position and
contigs are sorted by size. Red line shows p-value that identifies
outliers. Outlier SNPs above this threshold are highlighted and labeled.
Bold labels indicate the presence of at least one non-synonymous SNP in
that gene.}

@@

\caption{\label{fig-synteny-plot}Synteny plot showing conserved gene
order in \emph{Xenopus tropicalis} and \emph{Rana mucosa} for the
outlier region containing MHC Class I Classical and MHC Class III gene
regions. The plot was created with SimpleSynteny \citep{veltri2016}
using \emph{Xenopus tropicalis} Chromosome 8 (NC\_030684.2, genbank
accession GCA\_000004195.4) and \emph{Rana mucosa} Contig19. Asterices
indicate the location of SNP outliers in the TCF19 and VARS1 genes. Gap
sizes for each contig representation are labeled.}

@@

{\centering \includegraphics[width=0.60\textwidth]{figures/map_translocation_points.png}

@@

\caption{\label{fig-yosemap}Map showing the locations of translocated
and donor MYL frog populations in Yosemite National Park (park boundary
indicated by black polygon). Symbol shapes indicate the donor population
used for each translocation site. To obscure the exact locations of
populations, random noise was added to all point coordinates. Inset map
shows the location of Yosemite within California. In both maps,
elevation is indicated by the colored hillshade layer (dark green =
lowest elevation, white = highest elevation).}

@@

{\centering \includegraphics[width=0.60\textwidth]{figures/mcmc_areas_m2b.png}

@@

{\centering \includegraphics[width=0.60\textwidth]{figures/bdload_beforeafter.png}

@@

\caption{\label{fig-bdload-beforeafter}Evidence that frogs translocated
from recovering populations can maintain the benefits of
resistance/tolerance in non-natal habitats. For frogs translocated to
each of the 12 recipient sites, Bd loads for the period immediately
prior to translocation versus during the 1-year period after
translocation. Bd loads are expressed as the number of ITS1 copies per
skin swab, as estimated by qPCR of the Bd ITS1 region. Loads indicative
of severe disease are \textgreater{} 5.8 ITS copies (on a
log\textsubscript{10} scale; see SI Appendix text for details). Box plot
details are as in Fig. 4.}

@@

{\centering \includegraphics[width=0.60\textwidth]{figures/mcmc_areas_m1d.png}

@@

\caption{\label{fig-survival-postdens}Evidence that Bd load is not an
important predictor of post-translocation frog survival. For the best
model identifying predictors of 1-year frog survival following
translocation, estimated posterior density curves and shaded 95\%
uncertainty intervals for the intercept and all predictor variables. In
the Bayesian framework in which the model was developed, variables are
considered important predictors if the associated uncertainty interval
does not overlap zero (indicated by the dashed red line).}

@@

{\centering \includegraphics[width=0.60\textwidth]{figures/compare_surv_probs.jpg}

@@

\caption{\label{fig-compare_surv_probs}Evidence that frog
resistance/tolerance in non-natal habitats is maintained across
generations. Comparison of the average yearly adult survival
probabilities for adults translocated to each of 3 sites versus adults
that were naturally recruited at each site (as a result of reproduction
by translocated frogs). In contrast to
Fig. 3, these are not survival
probabilities from the first year following translocation, but instead
represent averaged survival probabilities across multiple years and
cohorts. Points are median estimates and error bars give the 95\%
uncertainty intervals around the estimates, accounting for yearly
variation in survival probabilities. The large error bars for 70550
reflect the large within and among individual variability in yearly
survival probability at this site (as also reflected in one-year
survival probabilities in Fig. 3). All
estimates are derived from the mrmr package.}

@@

{\centering \includegraphics[width=0.60\textwidth]{figures/pop_viability_figures_for_supp.jpg}

@@

\hypertarget{tbl-survival-earlylate}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0654}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2056}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2430}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2056}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2804}}@{}}
\caption{\label{tbl-survival-earlylate}Association between the
proportion of populations translocated early versus late in the study
period (\textless{} 2013 or $\geq$ 2013, respectively) and
probability of survival (\textless{} 0.5 or $\geq$ 0.5). Based on
the viability analysis, survival probabilities \textless{} 0.5 and
$\geq$ 0.5 produced 50-year extinction probabilities of 1 and
\textless{} 0.5, respectively.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Period
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Survival probability
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Number of translocations
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Total translocations
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Proportion of translocations
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Period
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Survival probability
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Number of translocations
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Total translocations
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Proportion of translocations
\end{minipage} \\
\midrule()
\endhead
early & \textless{} 0.5 & 4 & 5 & 0.800 \\
early & $\geq$ 0.5 & 1 & 5 & 0.200 \\
late & \textless{} 0.5 & 2 & 7 & 0.286 \\
late & $\geq$ 0.5 & 5 & 7 & 0.714 \\
\bottomrule()
\end{longtable}
